Enriching Object Detection with 2D-3D Registration\\and Continuous Viewpoint Estimation

### 2015-04-01 approach.tex
grammar mod

In this section, we describe our approach to generating 3D
exemplar-based templates on-the-fly. It is based on 3D CAD model
rendering (Sect.~\ref{sec:rendering}) followed by feature extraction
and applying a whitening procedure. Based on the original whitening
formulation (Sect.~\ref{sec:who}), we propose three novel extensions that
enable the application of whitening to the on-the-fly setting.
%
First, we adapt the whitening to the specific case of rendered
images (Sect.~\ref{sec:nzwho}). Second, we show how to speed up the
whitening by two orders of magnitudes for high-resolution templates
(Sect.~\ref{sec:fastwhiten}). And third, we improve the runtime of our
3D exemplar template detectors also at test time, by performing
convolutions in the frequency domain (Sect.~\ref{sec:fft}).

@@

We used an off-the-shelf rendering engine to generate a realistic rendering
and a depth map. The CAD models we used contain texture and material
information. These make the rendering more realistic and allows us to transfer
natural image statistics to rendered images. Along
with a rendering, we extracted a depth map which can be used for
reconstruction\ref{fig:rendering}.

@@

\subsection{Whitened Histograms of Orientations (WHO)}
\label{sec:who}
Our technique for rendering and generating exemplar template detectors
on-the-fly is drawing from recent work. Hariharan \etal. introduced Whitened
Histograms of Orientations (WHO), which uses feature statistics from natural
images to create a large number of classifiers analytically using Linear
Discriminant Analysis(LDA) rather than training SVM
classifiers\cite{Hariharan12}. The confidence score for data $x_i$ can be
defined as $ S(x_i) = w_{x_t}^T x_i$ where $w_{x_t} = \Sigma^{-1} (x_t - \mu)$
is an LDA classifier for a template $x_t$. Since collecting covariance matrices
for all possible template shape is intractable, \cite{Hariharan12} assumed
Wide-Sense Stationarity(WSS) of HOG features and generated a covariance
$\Sigma$ from autocovariance $\Gamma$ collected on a large collection of
natural images. i.e. for a 31 dim HOG feature $x_t$ at location $t$

\begin{align}
    E[x_t] & = \mu\\
    E[(x_{t} - \mu ) (x_{\tau} - \mu)] & = E[(x_0 - \mu)(x_{\tau - t} - \mu)]
\end{align}

@@

Alternatively, if the background is left textureless (see
Fig.~\ref{fig:rendering}), whitening the resulting HOG template
, strong negative weights are introduced in the textureless region (by
subtracting the mean $\mu$) as seen in Fig.~\ref{fig:whocomparison}.
This could result in positive matches being suppressed due to spurious
background gradients.

NZ-WHO removes these artifacts so that the background has no effect on the
template response. Let a vectorized HOG features of a rendering image $x = [
\begin{array}{ccccc}x_1^T & x_2^T & \cdots & x_n^T\end{array} ]^T \in
\mathcal{R}^{nd}$ where $x_i$ is the $i$th HOG cell feature and $n$ is the
number HOG cells and $d$ is the dimension of HOG feature. We create a new
vector $\bar{x}$ which contains only the non-zero HOG cell features of $x$. To be
specific, let $I_d\in \mathcal{R}^{d\times d}$ be the identity matrix, $\bar{n}$ be
the number of non-zero HOG cells, and a matrix $S\in \mathcal{R}^{\bar{n}d \times
    nd}$ be the masking matrix that selects non-zero HOG cells. For instance, a
template has  $n=3, \bar{n}=2$, and only the second HOG cell has $0$ norm, then

@@

The matrix selects HOG features that corresponds to the first cell and the third cell.

@@

\begin{figure}[t]
  \begin{center}
%   \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
    % include whitening all centered cells
    % \includegraphics[width=0.32\linewidth]{whiten_all_crop}
    \setlength\tabcolsep{3pt}
    \begin{tabular}{ccc}
      HOG & WHO & NZ-WHO \\
%     \begin{turn}{90}$w_+$\end{turn} &
    \includegraphics[width=0.28\linewidth]{hog_crop} &
    \includegraphics[width=0.28\linewidth]{whiten_all_crop} &
    \includegraphics[width=0.28\linewidth]{whiten_non_zero_crop} \\
    % include whitening all centered cells
     % \includegraphics[width=0.282\linewidth]{whiten_all_neg_crop} 
%     \begin{turn}{90}$-w_-$\end{turn} &
     \includegraphics[width=0.28\linewidth]{hog_neg_crop} &
     \includegraphics[width=0.28\linewidth]{whiten_all_neg_crop}  &
     \includegraphics[width=0.28\linewidth]{whiten_non_zero_neg} \\
    % include whitening all centered cells
 %    \begin{turn}{90}ihog\cite{vondrick2013}\end{turn} &
    % \cite{vondrick2013}& 
     \includegraphics[width=0.28\linewidth]{ihog_hog200_crop.png} &
     \includegraphics[width=0.28\linewidth]{ihog_whiten_all200_crop.png} &
     \includegraphics[width=0.28\linewidth]{ihog_whiten_non_zero200_crop.png} \\
 \end{tabular}
  \end{center}
  \caption{Comparison of HOG, WHO and NZ-WHO. Visualization of positive weights (first row),  visualization of negative weights (second row), HOGgles \cite{vondrick2013} (third row). Note that for WHO, whitening all cell result in strong negative edges on the empty region}
  \label{fig:whocomparison}
\end{figure}

@@

Synthesizing an LDA template Eq.~\ref{eq:nz-who} requires solving the system of
linear equations, $\bar{\Sigma} \bar{w} = (\bar{x} - \bar{\mu})$. In
\cite{Hariharan12}, the authors make use of the fact that covariance matrices
are symmetric and positive semidefinite to solve the system via the Cholesky
decomposition with Gaussian Elimination, which requires $O(n^3)$ time.

The Conjugate Gradient method is an iterative algorithm for solving symmetric
positive definite systems which runs in $O(n^2\kappa)$ time, where $\kappa$ is
the condition number of the matrix. \cite{Shewchuk94}.
%
This makes Conjugate Gradient faster than decomposition for matrices with small
condition numbers relative to their size.

The covariance matrix for HOG templates is typically
ill-conditioned\cite{Hariharan12}, but adding a small regularization constant to
the diagonal reduces its condition number.
We use a constant of $0.15$, which reduces the condition number from $10^{20}$
to $50$, much smaller than the dimension of the matrix (7000).

@@

We report the real time analysis of whitening using decomposition and
conjugate gradient methods in Fig.~\ref{fig:whotime}. (a) compares the
absolute runtime of the different methods while (b) gives the obtained
speedup. We see that % We compare the speed of each of template generation methods in
% Fig.~\cite{fig:covariancetime_crop} Using naive Cholesky decomposition,
WHO template takes several seconds for realistic template sizes of
several hundred cells, while using the iterative Conjugate Gradient
method, it only takes 60ms. If we use NZ-WHO, we can gain extra speed
up since we only whitens non-zero cells.

In addition, since the iterative Conjugate Gradient method directly
tries to reduce the residual (the norm of $y-Ax$ for $Ax = y$), it is
more numerically stable than Cholesky decomposition with Gaussian
Elimination. We vary the number of cells in a template and show that
the residual of NZ-WHO is smaller than WHO Fig.~\ref{fig:whoresidual}.
% Also, the residual from the Conjugate Gradient method (the norm of $y-Ax$ for $Ax = y$) is smaller than 
% that of Cholesky decomposition.

@@

We generate high resolution templates with more than 250 HOG cells to capture
details of an object to give accurate 2D-3D matching. These large templates
cause computational burden when computing convolution. Though good for
detecting accurate model and pose, convolution of high-resolution templates are
much slower since computation time scales linearly with the number of HOG
cells in the template. To overcome and speed up the convolution, we used FFT
based GPU convolution \cite{Podlozhnyuk} which scales. Briefly, for length $n$
signal and length $m$ filter, naive convolution takes $O(nm)$ time whereas
FFT-based convolution takes $O\left( (n + m)\log (n+m) \right)$ time. For large
$m$ (high resolution templates), we can gain computational advantage.

>>>>

In this section, we describe our approach to generating 3D
exemplar-based templates on-the-fly. It is based on 3D CAD model
rendering (Sect.~\ref{sec:rendering}) followed by feature extraction
and whitening. Based on the original whitening
formulation (Sect.~\ref{sec:who}), we propose three novel extensions that
enable the application of whitening to the on-the-fly setting.
%
First, we adapt the whitening to the specific case of rendered
images (Sect.~\ref{sec:nzwho}). Second, we show how to speed up the
whitening by two orders of magnitude for high-resolution templates
(Sect.~\ref{sec:fastwhiten}). Third, we improve the evaluation of our
3D exemplar template detectors at test time by performing
convolutions in the frequency domain (Sect.~\ref{sec:fft}).

@@

We used an off-the-shelf rendering engine to generate a realistic rendering
and a depth map. The CAD models we used contain texture and material
information. These make the rendering more realistic and allow us to transfer
natural image statistics to rendered images. During rendering we extract a depth map which can be used for
reconstruction\ref{fig:rendering}.

@@

\subsection{Whitened Histograms of Orientations (WHO)}
\label{sec:who}
Our technique for rendering and generating exemplar template detectors
on-the-fly draws from recent work by Hariharan \etal~\cite{Hariharan12}. They introduced Whitened
Histograms of Orientations (WHO), which uses feature statistics from natural
images to create a large number of classifiers analytically using Linear
Discriminant Analysis (LDA) rather than training SVM
classifiers. The confidence score for data $x_i$ can be
defined as $ S(x_i) = w_{x_t}^T x_i$ where $w_{x_t} = \Sigma^{-1} (x_t - \mu)$
is an LDA classifier for a template $x_t$. Since collecting covariance matrices
for all possible template shape is intractable, \cite{Hariharan12} assumed
Wide-Sense Stationarity(WSS) of HOG features and generated a covariance
$\Sigma$ from autocovariance $\Gamma$ collected on a large collection of
natural images. For a 31 dim HOG feature $x_t$ at location $t$

\begin{align}
    E[x_t] & = \mu\\
    E[(x_{t} - \mu ) (x_{\tau} - \mu)] & = E[(x_0 - \mu)(x_{\tau - t} - \mu)].
\end{align}

@@

Alternatively, if the background is left textureless (see
Fig.~\ref{fig:rendering}), whitening the resulting HOG template introduces strong negative weights in the textureless region (by
subtracting the mean $\mu$), as seen in Fig.~\ref{fig:whocomparison}.
This could result in positive matches being suppressed due to spurious
background gradients.

NZ-WHO removes these artifacts so that the background has no effect on the
template response. Let a vectorized HOG features of a rendering image $x = [
\begin{array}{ccccc}x_1^T & x_2^T & \cdots & x_n^T\end{array} ]^T \in
\mathcal{R}^{nd}$ where $x_i$ is the $i$th HOG cell feature, $n$ is the
number of HOG cells and $d$ is the dimension of the HOG feature. We create a new
vector $\bar{x}$ which contains only the non-zero HOG cell features of $x$. To be
specific, let $I_d\in \mathcal{R}^{d\times d}$ be the identity matrix, $\bar{n}$ be
the number of non-zero HOG cells, and a matrix $S\in \mathcal{R}^{\bar{n}d \times
    nd}$ be the masking matrix that selects non-zero HOG cells. For instance, a
template has  $n=3, \bar{n}=2$, and only the second HOG cell has $0$ norm, then

@@

The matrix selects HOG features that correspond to the first and third cells.

@@

\begin{figure}[t]
  \begin{center}
%   \fbox{\rule{0pt}{2in} \rule{0.9\linewidth}{0pt}}
    % include whitening all centered cells
    % \includegraphics[width=0.32\linewidth]{whiten_all_crop}
    \setlength\tabcolsep{3pt}
    \begin{tabular}{ccc}
      HOG & WHO & NZ-WHO \\
%     \begin{turn}{90}$w_+$\end{turn} &
    \includegraphics[width=0.28\linewidth]{hog_crop} &
    \includegraphics[width=0.28\linewidth]{whiten_all_crop} &
    \includegraphics[width=0.28\linewidth]{whiten_non_zero_crop} \\
    % include whitening all centered cells
     % \includegraphics[width=0.282\linewidth]{whiten_all_neg_crop} 
%     \begin{turn}{90}$-w_-$\end{turn} &
     \includegraphics[width=0.28\linewidth]{hog_neg_crop} &
     \includegraphics[width=0.28\linewidth]{whiten_all_neg_crop}  &
     \includegraphics[width=0.28\linewidth]{whiten_non_zero_neg} \\
    % include whitening all centered cells
 %    \begin{turn}{90}ihog\cite{vondrick2013}\end{turn} &
    % \cite{vondrick2013}& 
     \includegraphics[width=0.28\linewidth]{ihog_hog200_crop.png} &
     \includegraphics[width=0.28\linewidth]{ihog_whiten_all200_crop.png} &
     \includegraphics[width=0.28\linewidth]{ihog_whiten_non_zero200_crop.png} \\
 \end{tabular}
  \end{center}
  \caption{Comparison of HOG, WHO and NZ-WHO. Visualization of positive weights (first row),  visualization of negative weights (second row), HOGgles \cite{vondrick2013} (third row). Note that for WHO, whitening all cells results in strong negative edges on the empty region}
  \label{fig:whocomparison}
\end{figure}

@@

Synthesizing the LDA template in Eq.~\ref{eq:nz-who} requires solving the system of
linear equations, $\bar{\Sigma} \bar{w} = (\bar{x} - \bar{\mu})$. In
\cite{Hariharan12}, the authors make use of the fact that covariance matrices
are symmetric and positive semidefinite to solve the system via the Cholesky
decomposition with Gaussian Elimination, which requires $O(n^3)$ time.

The Conjugate Gradient method is an iterative algorithm for solving symmetric
positive definite systems which runs in $O(n^2\kappa)$ time, where $\kappa$ is
the condition number of the matrix~\cite{Shewchuk94}.
%
This makes Conjugate Gradient faster than decomposition for matrices with small
condition numbers relative to their size.

The covariance matrix for HOG templates is typically
ill-conditioned~\cite{Hariharan12}, but adding a small regularization constant to
the diagonal reduces its condition number.
We use a constant of $0.15$, which reduces the condition number from $10^{20}$
to $50$, much smaller than the dimension of the matrix (7000).

@@

We report the real time analysis of whitening using decomposition and
Conjugate Gradient methods in Fig.~\ref{fig:whotime}. (a) compares the
absolute runtime of the different methods while (b) gives the obtained
speedup. We see that % We compare the speed of each of template generation methods in
% Fig.~\cite{fig:covariancetime_crop} Using naive Cholesky decomposition,
whitening the WHO template takes several seconds for realistic template sizes of
several hundred cells, but only 60ms using the Conjugate Gradient
method. If we use NZ-WHO, we can gain extra speed
up since we only whiten non-zero cells.

In addition, since the iterative Conjugate Gradient method directly
tries to reduce the residual (the norm of $y-Ax$ for $Ax = y$), it is
more numerically stable than Cholesky decomposition with Gaussian
Elimination. In Fig.~\ref{fig:whoresidual} we vary the number of cells in a template and show that
the residual of NZ-WHO is smaller than that of WHO.
% Also, the residual from the Conjugate Gradient method (the norm of $y-Ax$ for $Ax = y$) is smaller than 
% that of Cholesky decomposition.

@@

We generate high resolution templates with more than 250 HOG cells to capture
details of an object to give accurate 2D-3D matching. These large templates
cause computational burden when computing convolution. Though good for
accurately determining model and pose, these large templates slow down the convolution since computation time scales linearly with the number of HOG
cells in the template. To overcome this, we used FFT-based GPU convolution \cite{Podlozhnyuk}. Briefly, for length $n$
signal and length $m$ filter, naive convolution takes $O(nm)$ time whereas
FFT-based convolution takes $O\left( (n + m)\log (n+m) \right)$ time. For large
$m$ (high resolution templates), we can gain computational advantage.

### 2015-04-01 egpaper_final.tex
grammar mod

%%%%%%%%% ABSTRACT
\begin{abstract}
A large body of recent work on object detection has focused on exploiting 3D CAD
model databases to improve detection performance. Many of these approaches work
by aligning exact 3D models to images using templates generated from renderings
of the 3D models at a finite set of discrete viewpoints. The training procedures
for these approaches, however, are very expensive and require gigabytes of
memory and storage, and the viewpoint discretization hampers pose estimation
performance.

We propose an efficient method for synthesizing templates from 3D models that
runs on the fly -- that is, it quickly produces detectors for an arbitrary
viewpoint of a 3D model without expensive dataset-dependent training or template
storage. Given a 3D model and an arbitrary continuous detection viewpoint, our
method synthesizes a discriminative template by extracting features from a
rendered view of the object and decorrelating spatial dependences among the
features. Our decorrelation procedure relies on a gradient-based algorithm that
is more numerically stable than standard decomposition-based procedures, and we
efficiently search for candidate detections by computing FFT-based template
convolutions. Due to the speed of our template synthesis procedure, we are able
to perform joint optimization on scale, translation, continuous rotation, and
focal length. We provide an efficient GPU implementation of our algorithm, and
we validate its performance on 3D Object Classes and PASCAL3D+ datasets.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
A large body of recent work on object detection has focused on exploiting 3D CAD
model databases to improve detection performance. Many of these approaches work
by aligning exact 3D models to images using templates generated from renderings
of the 3D models at a set of discrete viewpoints. However, the training procedures for these approaches are very expensive and require gigabytes of
memory and storage, while the viewpoint discretization hampers pose estimation
performance.

We propose an efficient method for synthesizing templates from 3D models that
runs on the fly -- that is, it quickly produces detectors for an arbitrary
viewpoint of a 3D model without expensive dataset-dependent training or template
storage. Given a 3D model and an arbitrary continuous detection viewpoint, our
method synthesizes a discriminative template by extracting features from a
rendered view of the object and decorrelating spatial dependences among the
features. Our decorrelation procedure relies on a gradient-based algorithm that
is more numerically stable than standard decomposition-based procedures, and we
efficiently search for candidate detections by computing FFT-based template
convolutions. Due to the speed of our template synthesis procedure, we are able
to perform joint optimization of scale, translation, continuous rotation, and
focal length. We provide an efficient GPU implementation of our algorithm, and
we validate its performance on 3D Object Classes and PASCAL3D+ datasets.

### 2015-04-01 experiments.tex
Added selector matrix $S$ that generates $x', \mu', \Sigma'$.
Major approach modification don. minor mcmc and overview
Approach major whiten to LDA update
MCMC minor math update
experiment, major PASCAL12 table to scriptsize
grammar mod

%
In this section, we give an experimental evaluation of our approach,
highlighting three different aspects.
%
First, we verify that our NZ-WHO method delivers performance that is
at least on par with the original WHO formulation~\cite{Hariharan12} in terms of
accuracy, while at the same time resulting in large computational
savings (Sect.~\ref{sec:variations}).
%
Second, we demonstrate that our method can be used for multi-view
object class detection in isolation. It can be applied in a sliding
window fashion and deliver 2D bounding box as well as viewpoint
information. Our method compares on par to state-of-the-art in this
case (Sect.~\ref{sec:exp_iso}).
%
And third, we show that our method can be used to complement the
detections provided by an existing object class detector, such
as DPM~\cite{Felzenszwalb10} or RCNN~\cite{Girshick14}. In this case,
we show a considerable performance improvement compared to previous
work in the task of joint object class detection and VP
estimation (Sect.~\ref{sec:exp_enrich}).

@@

\subsection{WHO Variants}
\label{sec:variations}
To validate our approach, we run an ensemble of NZ-WHO templates as a
bank of exemplar detectors and compare its performance with other WHO
variants, notable WHO~\cite{Hariharan12} and the original non-whitened
HOG~\cite{Dalal05}. In addition, we evaluate WHO-CG and WHO-CG-Z:
WHO-CG uses iterative Conjugate Gradient method to generate
WHO. WHO-CG-Z whitens the whole template, but zeros out textureless
region. NZ-WHO-CG is the NZ-WHO which whitens only non-zero cells using
iterative Conjugate Gradients  (our method).

@@

To that end, we create an ensemble of NZ-WHO templates (Sect.~\ref{sec:approach})
using $9$ different CAD models and a total of 192 different
viewpoints, using 4 different elevation angles, 24 azimuth angles,
and 2 focal lengths. We run the entire ensemble exhaustively over
each test image, in a sliding window fashion.

@@

\paragraph{Results.}
Tab.~\ref{tab:3dobject} gives the corresponding results, comparing our
method to two recent state-of-the-art baselines, the aspect layout
model (ALM~\cite{Xiang12}), and the
DPM-VOC+VP~\cite{Pepik12}. Fig.~\ref{fig:3dobject} gives
qualitative results.

@@

To show such ability, we evaluate our method on the PASCAL3D+
dataset~\cite{Xiang14}. This dataset augments PASCAL 2012 images with
high quality viewpoint annotations thus is ideal to measure pose
estimation. The dataset proposes new metric called Average Viewpoint
Precision (AVP) where it measures the area under viewpoint precision
and detection recall curve. The viewpoints is measured by azimuth
similarity. If the distance between predicted azimuth and ground truth
azimuth is below a certain threshold, the viewpoint is correct.
The baseline methods V-DPM~\cite{Xiang14} and
DPM-VOC+VP~\cite{Pepik12}) reported on the PASCAL3D+ dataset are
variants of DPM~\cite{Felzenszwalb10} where each component of DPM
account for various azimuths. Thus V-DPM and VOC-DPM provide discrete
azimuth only whereas our method provides 3D viewpoint (yaw, pitch,
roll), CAD instance (model index, rendering, depth) and focal length.
%
\begin{figure}[t]
 \begin{center}
    \setlength\tabcolsep{0pt}
    \begin{tabular}{ccc}
    % \includegraphics[width=0.9\linewidth]{tuning} 
   \includegraphics[width=0.33\linewidth]{tuning/1.png} &
   \includegraphics[width=0.33\linewidth]{tuning/2.png} &
   \includegraphics[width=0.33\linewidth]{tuning/3.png} \\[-5pt]
%   (a) & (b) & (c)\\
   \includegraphics[width=0.33\linewidth]{tuning/4.png} &
   \includegraphics[width=0.33\linewidth]{tuning/5.png} &
   \includegraphics[width=0.33\linewidth]{tuning/6.png} \\[-5pt]
%   (d) & (e) & (f)\\
   \end{tabular}
 \end{center}
 \caption{Effect of fine tuning. (left) original image, (middle) initial detection, (right) continuous fine tuning using Single-Component Metropolis Hastings}
 \label{fig:tuning}
\end{figure}

\paragraph{Setup.}
We use detection bounding boxes provided by both object detectors and
augment it using our method to give viewpoint estimation. We use
detection bounding boxes from two different methods:
DPM-VOC+VP\cite{Pepik12} and R-CNN~\cite{Girshick14}, both in their
variants trained on 8 viewpoint categories, since these perform best
in terms of AP. For both cases,
we use the original score for plotting precision-recall curves
(meaning that we can not improve over their AP), but
enrich the detections with additional viewpoint information provided
by our method. We fall back to a default
viewpoint prediction ($0$ angles) in case the confidence of our
method falls below a threshold.

@@

Quantitative results are given in Tab.~\ref{tab:pascal12} in terms of
AP and AVP and in Fig.~\ref{fig:car_cnn_ap} in terms of
precision-recall plots and viewpoint confusion matrices. Example
outputs of our method applied to candidate object detections are given
in Fig.~\ref{fig:pascal12cnn}, and the effect of fine-tuning is
visualized in Fig.~\ref{fig:tuning}. More qualitative results are part of
supplemental material
\paragraph{Results.}
%
In Tab.~\ref{tab:pascal12}, we make mainly two observation. First, we
see that adding our method to DPM-VOC+VP consistently improves
performance in terms of AVP, for both car and bicycle, and across all
viewpoint bins. This is already the case for our discrete method: the
improvement ranges from $0.3$ for bicycle-8v to $5.8$ for car-4v. 
%
Using the R-CNN as the base detector increases AVP even more, in
particular for the bicycle class: for bicycle-4v, our discrete method
improves over the corresponding DPM-VOC+VP result by $12.6$.

The second observation is that the fine-tuning based on MCMC can
indeed improve pose estimation performance slightly, e.g., by $1.1$
for bicycle-16v and using the DPM-VOC+VP as the base detector, or by
$1.1$ for bicycle-16v when using the R-CNN as the base detector.
%
In Fig.~\ref{fig:tuning}, we visualize the effect of fine-tuning
qualitatively for two different test images. In both cases, the
fine-tuned visually comes closer to the true 3D pose.
% We got significant improvement on viewpoint estimation by using our
% method on top of general purpose detectors. Note that our method
% estimates azimuth, elevation, in-plane rotation, focal length
% estimation, fine-grained categorization whereas V-DPM, VOC-DPM only
% estimate discrete viewpoints.

@@

  \includegraphics[width=0.24\textwidth]{car_cnn/2a.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/2b.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/2c.png} &   
  \includegraphics[width=0.24\textwidth]{car_cnn/2e.png}  \\  

@@

\caption{Example enriched bounding boxes. Given R-CNN\cite{Girshick14}

@@

  reasonably. On the first column, R-CNN detection bounding boxes
  overlaid. From the second to the last columns, output of our method
  given a R-CNN box. Blue boxes are R-CNN output and purple boxes are

@@

    \footnotesize

@@

@@

@@

\caption{Average Precision(AP) and Average Viewpoint Precision (AVP) on

@@

>>>>

%
In this section, we give an experimental evaluation of our approach,
highlighting three different aspects.
%
First, we verify that our NZ-WHO method delivers performance that is
at least on par with the original WHO formulation~\cite{Hariharan12} in terms of
accuracy, while at the same time resulting in large computational
savings (Sect.~\ref{sec:variations}).
%
Second, we demonstrate that our method can be used for multi-view
object class detection in isolation. It can be applied in a sliding
window fashion and deliver 2D bounding box as well as viewpoint
information. Our method is competitive with the state-of-the-art in this
case (Sect.~\ref{sec:exp_iso}).
%
Finally, we show that our method can be used to complement the
detections provided by an existing object class detector, such
as DPM~\cite{Felzenszwalb10} or RCNN~\cite{Girshick14}. In this case,
we show a considerable performance improvement compared to previous
work in the task of joint object class detection and VP
estimation (Sect.~\ref{sec:exp_enrich}).

@@

\subsection{WHO Variants}
\label{sec:variations}
To validate our approach, we run an ensemble of NZ-WHO templates as a
bank of detectors and compare its performance with other WHO
variants, notable WHO~\cite{Hariharan12} and the original non-whitened
HOG~\cite{Dalal05}. In addition, we evaluate WHO-CG and WHO-CG-Z:
WHO-CG uses iterative Conjugate Gradient method to generate
WHO. WHO-CG-Z whitens the whole template, but zeros out textureless
region. NZ-WHO-CG is the NZ-WHO which whitens only non-zero cells using
iterative Conjugate Gradients  (our method).

@@

To that end, we create an ensemble of NZ-WHO templates (Sect.~\ref{sec:approach})
using $9$ different CAD models and a total of 192 different
viewpoints: 4 elevation angles, 24 azimuth angles,
and 2 focal lengths. We run the entire ensemble exhaustively over
each test image in a sliding window fashion.

@@

\paragraph{Results.}
Tab.~\ref{tab:3dobject} gives the corresponding results, comparing our
method to two recent state-of-the-art baselines, the aspect layout
model (ALM~\cite{Xiang12}), and the
DPM-VOC+VP~\cite{Pepik12}. Fig.~\ref{fig:3dobject_vis} gives
qualitative results.

@@

To show such ability, we evaluate our method on the PASCAL3D+
dataset~\cite{Xiang14}. This dataset augments PASCAL 2012 images with
high quality viewpoint annotations thus is ideal to measure pose
estimation. The dataset proposes new metric called Average Viewpoint
Precision (AVP) where it measures the area under viewpoint precision
and detection recall curve. The viewpoints is measured by azimuth
similarity. If the distance between predicted azimuth and ground truth
azimuth is below a certain threshold, the viewpoint is correct.
The baseline methods V-DPM~\cite{Xiang14} and
DPM-VOC+VP~\cite{Pepik12}) reported on the PASCAL3D+ dataset are
variants of DPM~\cite{Felzenszwalb10} where each component of DPM
account for various azimuths. Thus V-DPM and VOC-DPM provide discrete
azimuths only whereas our method provides 3D viewpoint (yaw, pitch,
roll), CAD instance (model index, rendering, depth) and focal length.
%
\begin{figure}[t]
 \begin{center}
    \setlength\tabcolsep{0pt}
    \begin{tabular}{ccc}
    % \includegraphics[width=0.9\linewidth]{tuning} 
   \includegraphics[width=0.33\linewidth]{tuning/1.png} &
   \includegraphics[width=0.33\linewidth]{tuning/2.png} &
   \includegraphics[width=0.33\linewidth]{tuning/3.png} \\[-5pt]
%   (a) & (b) & (c)\\
   \includegraphics[width=0.33\linewidth]{tuning/4.png} &
   \includegraphics[width=0.33\linewidth]{tuning/5.png} &
   \includegraphics[width=0.33\linewidth]{tuning/6.png} \\[-5pt]
%   (d) & (e) & (f)\\
   \end{tabular}
 \end{center}
 \caption{Effect of fine tuning. (left) original image, (middle) initial detection, (right) continuous fine tuning using Single-Component Metropolis Hastings}
 \label{fig:tuning}
\end{figure}

\paragraph{Setup.}
We use detection bounding boxes provided by both object detectors and
augment use our method to perform fine-grained viewpoint estimation. We use
detection bounding boxes from two different methods:
DPM-VOC+VP\cite{Pepik12} and R-CNN~\cite{Girshick14}, both in their
variants trained on 8 viewpoint categories, since these perform best
in terms of AP. For both cases,
we use the original score for plotting precision-recall curves
(meaning that we can not improve over their AP). We revert back to a default
viewpoint prediction ($0$ angles) in case the confidence of our
method falls below a threshold.

@@

Quantitative results are given in Tab.~\ref{tab:pascal12} in terms of
AP and AVP and in Fig.~\ref{fig:car_cnn_ap} in terms of
precision-recall plots and viewpoint confusion matrices. Example
outputs of our method applied to candidate object detections are given
in Fig.~\ref{fig:pascal12cnn}, and the effect of fine-tuning is
visualized in Fig.~\ref{fig:tuning}. More qualitative results can be found in the
supplemental material.
\paragraph{Results.}
%
In Tab.~\ref{tab:pascal12}, we make two main observations. First, we
see that adding our method to DPM-VOC+VP consistently improves
performance in terms of AVP, for both car and bicycle, and across all
viewpoint bins. This is already the case for our discrete method: the
improvement ranges from $0.3$ for bicycle-8v to $5.8$ for car-4v. 
%
Using the R-CNN as the base detector increases AVP even more, in
particular for the bicycle class: for bicycle-4v, our discrete method
improves over the corresponding DPM-VOC+VP result by $12.6$.

The second observation is that the fine-tuning based on MCMC can
indeed improve pose estimation performance slightly, e.g., by $1.1$
for bicycle-16v and using the DPM-VOC+VP as the base detector, or by
$1.1$ for bicycle-16v when using the R-CNN as the base detector.
%
In Fig.~\ref{fig:tuning}, we visualize the effect of fine-tuning
qualitatively for two different test images. In both cases, the
fine-tuned pose is a better visual match to the true 3D pose.
% We got significant improvement on viewpoint estimation by using our
% method on top of general purpose detectors. Note that our method
% estimates azimuth, elevation, in-plane rotation, focal length
% estimation, fine-grained categorization whereas V-DPM, VOC-DPM only
% estimate discrete viewpoints.

@@

%  \includegraphics[width=0.24\textwidth]{car_cnn/2a.png} &   
%  \includegraphics[width=0.24\textwidth]{car_cnn/2b.png} &   
%  \includegraphics[width=0.24\textwidth]{car_cnn/2c.png} &   
%  \includegraphics[width=0.24\textwidth]{car_cnn/2e.png}  \\  

@@

\caption{Example enriched bounding boxes. Given R-CNN~\cite{Girshick14}

@@

  reasonably. The first column shows bounding boxed produced by R-CNN detection. Subsequent columns show the output of our method given this bounding box input. Blue boxes are R-CNN output and purple boxes are

@@

    \scriptsize

@@

    aeroplane-4v        & 40.0 / 34.6         & 41.5 / 37.4             & 41.1 / 32.7       & 41.1 / 30.5        & 67.8 / \textbf{41.3}         & 67.8 / 40.8 \\ \hline
    aeroplane-8v        & 39.8 / 23.4         & 40.5 / 28.6             & 41.1 / 26.8       & 41.1 / 24.2        & 67.8 / 26.8         & 67.8 / \textbf{27.4} \\ \hline
    aeroplane-16v       & 43.6 / 15.4         & 38.0 / 15.9             & 41.1 / 17.7       & 41.1 / 16.3        & 67.8 / 21.2         & 67.8 / \textbf{21.5} \\ \hline
    aeroplane-24v       & 42.2 / 8.0          & 36.0 /  9.7             & 41.1 / 10.9       & 41.1 / 10.2        & 67.8 / 15.6         & 67.8 / \textbf{15.7} \\ \hline

@@

    chair-4v            & 11.1 / 6.8          & 8.7  / 6.1              &  11.4 / 7.1       & 11.4 / 6.7         & 25.2 / 10.6         & 25.2 / \textbf{10.7} \\ \hline
    chair-8v            & 11.4 / 5.9          & 11.3 / 9.4              &  11.4 / 6.6       & 11.4 / 6.6         & 25.2 / \textbf{9.4}          & 25.2 / 9.3 \\ \hline
    chair-16v           & 12.8 / 6.0          & 10.2 / 6.1              &  11.4 / 4.7       & 11.4 / 4.7         & 25.2 / \textbf{6.7}          & 25.2 / 6.7 \\ \hline
    chair-24v           & 12.6 / 4.4          & 8.0  / 4.2              &  11.4 / 3.1       & 11.4 / 3.4         & 25.2 / 4.6          & 25.2 / \textbf{5.4} \\ \hline

@@

\caption{Average Precision (AP) and Average Viewpoint Precision (AVP) on

@@

\begin{comment}
\hline
boat-4v             & 3.0 / 1.5           & 0.5 / 0.3               &  /        &  /         & 27.9 / 11.3         & 27.9 / 11.5 \\ \hline
boat-8v             & 5.8 / 1.0           & 0.5 / 0.2               &  /        &  /         & 27.9 / 5.4          & 27.9 / 5.4 \\ \hline
boat-16v            & 6.2 / 0.5           & 0.7 / 0.3               &  /        &  /         & 27.9 / 2.0          & 27.9 / 2.0 \\ \hline
boat-24v            & 6.0 / 0.3           & 5.3 / 2.2               &  /        &  /         & 27.9 / 1.8          & 27.9 / 2.5 \\ \hline
\end{comment}
